---
title: "Liberal Analysis"
author: "Nicolas Restrepo"
date: "11/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminary Analysis

```{r}
# Packages 

library(tidyverse)
library(knitr)
library(kableExtra)
theme_set(theme_light())
```

I am going to begin by loading in the data. 

```{r}

# Load the data in 

data <- read_csv("liberals_first.csv")

# Get rid of the first two rows which are uninformative 

data <- data[-(1:2), ]

# Load in data from the first batch 

p_data <- read_csv("pretest.csv")

p_data <- p_data[-(1:3), ]

# Now bind them together 

new_data <- rbind(data, p_data)

```

### Attention checks 

First, it would be important to see who passed the attention checks.

```{r}

# Get attention checks for the data 

new_data <- new_data %>% 
  mutate(passed_1 = ifelse(ach_1_1 == "extremely harmful", 1, 0), 
         passed_2 = ifelse(ach_2_1 == "not at all immoral", 1, 0), 
         passed_3 = ifelse(ach_3_1 == "moderately unexpected", 1, 0), 
         passed_4 = ifelse(ach_4_1 == "not at all harmful", 1, 0), 
         passed_5 = ifelse(ach_5_1 == "extremely immoral", 1, 0), 
         passed_6 = ifelse(ach_6_1 == "extremely unexpected", 1, 0), 
         sum_pass = passed_1 + passed_2 + passed_3 + passed_4 + passed_5 + passed_6)

# Now drop cases which did not pass the attention checks 

new_data <- new_data %>% 
  filter(sum_pass == 6)

```

### Checking the Pre-Screening

For the purposes of this test, I told Prolific I only wanted liberal respondents. We can check whether the pre-screening did a good job. 

```{r}

new_data <- new_data %>% 
  mutate(liberal = ifelse(ideology_1 == 1 | ideology_1 == 2 | ideology_1 == 3, 1, 0)) 

# Drop participants who are not liberal 

new_data <- new_data %>% 
  filter(liberal == 1)

```

### Analyzing the Data 

I am going to recode the data so we get numbers instead of strings:

```{r}

# Mutate all main questions 

data_recoded <- new_data %>% 
  mutate_at(vars(contains("_h_1")), funs(recode(., 
                                   "not at all harmful" = 1, 
                                   "slightly harmful" = 2, 
                                   "moderately harmful" = 3, 
                                   "very harmful" = 4, 
                                   "extremely harmful" = 5))) %>% 
  mutate_at(vars(contains("_i_1")), funs(recode(., 
                                   "not at all immoral" = 1, 
                                   "slightly immoral" = 2, 
                                   "moderately immoral" = 3, 
                                   "very immoral" = 4, 
                                   "extremely immoral" = 5))) %>% 
  mutate_at(vars(contains("_u_1")), funs(recode(., 
                                   "not at all unexpected" = 1, 
                                   "slightly unexpected" = 2, 
                                   "moderately unexpected" = 3, 
                                   "very unexpected" = 4, 
                                   "extremely unexpected" = 5)))
  
  
```

Now, I am going to focus solely on the responses to events:

```{r}

# Get of of the first set of uninformative columns 

data_recoded <- data_recoded %>% 
  select(-contains("ach")) %>% 
  select(pkp_h_1:msas_u_1)

# Now let's create a more amenable dataset 

d <- data_recoded %>% 
  mutate(id = 1:n()) %>% 
  select(id, everything()) 

# Rename all variables 

d <- d %>%  
  rename_all(
    funs(str_remove(., "_1"))
    ) %>% 
  mutate_if(is.character, as.numeric)

```

Now I am going to create a summary table for each question 

```{r}

# create a summary table for the values of each question 

gd <- d %>% 
  select(-id) %>% 
  gather(key = "scenario", 
         value = "score") %>% 
  group_by(scenario) %>% 
  summarise_all(funs(avg = mean(.), maximum = max(.), minimum = min(.), st_dv = sd(.))) %>% 
  mutate(ci_lower = avg - 1.96*st_dv, 
         ci_upper = avg + 1.96*st_dv) 

gd <- gd %>% 
  mutate(scenario = str_replace(scenario, "acr", "athlete_cheats_rival"), 
         scenario = str_replace(scenario, "adr", "athlete_deceives_referee"), 
         scenario = str_replace(scenario, "ayc", "athlete_yells_at_coach"), 
         scenario = str_replace(scenario, "ccr", "coach_cheers_rival"), 
         scenario = str_replace(scenario, "ddf", "daughter_disobeys_father"),
         scenario = str_replace(scenario, "ecc", "employee_conspires_with_comp."),
         scenario = str_replace(scenario, "elb", "employee_lies_to_boss"), 
         scenario = str_replace(scenario, "git", "girl_interrupts_teacher"),
         scenario = str_replace(scenario, "idb", "intern_disobeys_boss"), 
         scenario = str_replace(scenario, "jbd", "judge_befriends_defendant"), 
         scenario = str_replace(scenario, "mbr", "man_betrays_relative"), 
         scenario = str_replace(scenario, "mbw", "man_betrays_wife"), 
         scenario = str_replace(scenario, "mmc", "man_marries_cousin"),
         scenario = str_replace(scenario, "mmls", "man_makes_love_sister"), 
         scenario = str_replace(scenario, "mpsa", "married_has_sex_adulterer"), 
         scenario = str_replace(scenario, "msas", "mother_sexually_arouses_son"), 
         scenario = str_replace(scenario, "msn", "mayor_slanders_neighbor"), 
         scenario = str_replace(scenario, "pch", "person_hurts_child"), 
         scenario = str_replace(scenario, "php", "person_hires_prostitute"), 
         scenario = str_replace(scenario, "pkp", "person_kills_person"),
         scenario = str_replace(scenario, "scc", "student_cheats_classmate"), 
         scenario = str_replace(scenario, "sip", "student_insults_professor"), 
         scenario = str_replace(scenario, "thls", "teacher_hits_lazy_student"), 
         scenario = str_replace(scenario, "whh", "wife_hits_husband"), 
         scenario = str_replace(scenario, "tmdp", "teenager_mocks_disabled"))
  
```


Now, we are in a position to start making some summary plots 

```{r}

# Keep only harm questions 

gdh <- gd[str_which(gd$scenario, "_h$"),]

# Plot harm questions 

gdh %>% 
  mutate(scenario = str_remove(scenario, "_h$")) %>% 
  ggplot(aes(x = reorder(scenario, avg), y = avg)) + 
  geom_point(col = "darkred", fill = "darkred") + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), pch = 0) +
  coord_flip() + 
  labs(x = "", 
       y = "Ratings", 
       title = "Perceptions of Harmfulness")

# Now lets look at immorality 

gdi <- gd[str_which(gd$scenario, "_i$"),]

# Plot immorality

gdi %>% 
  mutate(scenario = str_remove(scenario, "_i$")) %>% 
  ggplot(aes(x = reorder(scenario, avg), y = avg)) + 
  geom_point(col = "darkred", fill = "darkred") + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), pch = 0) +
  coord_flip() + 
  labs(x = "", 
       y = "Ratings", 
       title = "Perceptions of Immorality")

# Lastly let's look at unexpectedness 

gdu <- gd[str_which(gd$scenario, "_u$"),]

# Plot Unexpectedness

gdu %>% 
  mutate(scenario = str_remove(scenario, "_u$")) %>% 
  ggplot(aes(x = reorder(scenario, avg), y = avg)) + 
  geom_point(col = "darkred", fill = "darkred") + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), pch = 0) +
  coord_flip() + 
  labs(x = "", 
       y = "Ratings", 
       title = "Perceptions of Unexpectedness")

```


