---
title: "Liberal Analysis"
author: "Nicolas Restrepo"
date: "11/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminary analysis of the liberal data 

```{r}
# Packages 
library(tidyverse)
library(knitr)
library(kableExtra)
theme_set(theme_light())
```

### Data Preparation 

I am going to begin by loading the data. We get two different data sources so we need to merge them.  

```{r}
# Load the data in 

second_data <- read_csv("liberals_complete.csv")

# Get rid of the first two rows which are uninformative 

second_data <- second_data[-(1:2), ]

# Load in data from the first batch 

p_data <- read_csv("pretest.csv")

p_data <- p_data[-(1:3), ]

# Now bind them together 

data <- rbind(second_data, p_data)

```

Everything seems okay. We have 105 observations but this is before we filter for people who are not liberal and who did not pass the attention checks. 

Let's begin by filtering the respondents who fail the attention checks. 

```{r}

# Get attention checks for the data 

data <- data %>% 
  mutate(passed_1 = ifelse(ach_1_1 == "extremely harmful", 1, 0), 
         passed_2 = ifelse(ach_2_1 == "not at all immoral", 1, 0), 
         passed_3 = ifelse(ach_3_1 == "moderately unexpected", 1, 0), 
         passed_4 = ifelse(ach_4_1 == "not at all harmful", 1, 0), 
         passed_5 = ifelse(ach_5_1 == "extremely immoral", 1, 0), 
         passed_6 = ifelse(ach_6_1 == "extremely unexpected", 1, 0), 
         sum_pass = passed_1 + passed_2 + passed_3 + passed_4 + passed_5 + passed_6)


# Now drop cases which did not pass the attention checks 

data <- data %>% 
  filter(sum_pass == 6)

```

Now, let's make sure we only have liberals in the sample.

```{r}
# Create variable that indicates if someone is liberal

data <- data %>% 
  mutate(liberal = ifelse(ideology_1 == 1 | ideology_1 == 2 | ideology_1 == 3, 1, 0)) 

# Drop participants who are not liberal 

data <- data %>% 
  filter(liberal == 1)

```

In total, we have 99 usable observations. 

I am also going to recode the data so we get numbers instead of strings. 

```{r}

# Mutate all main questions so that we get numbers 

data <- data %>% 
  mutate_at(vars(contains("_h_1")), funs(recode(., 
                                   "not at all harmful" = 1, 
                                   "slightly harmful" = 2, 
                                   "moderately harmful" = 3, 
                                   "very harmful" = 4, 
                                   "extremely harmful" = 5))) %>% 
  mutate_at(vars(contains("_i_1")), funs(recode(., 
                                   "not at all immoral" = 1, 
                                   "slightly immoral" = 2, 
                                   "moderately immoral" = 3, 
                                   "very immoral" = 4, 
                                   "extremely immoral" = 5))) %>% 
  mutate_at(vars(contains("_u_1")), funs(recode(., 
                                   "not at all unexpected" = 1, 
                                   "slightly unexpected" = 2, 
                                   "moderately unexpected" = 3, 
                                   "very unexpected" = 4, 
                                   "extremely unexpected" = 5)))
  
  
```

For now, I am going to select only the responses to events:

```{r}

# Get of of the first set of uninformative columns 

d <- data %>% 
  select(-contains("ach")) %>% 
  select(pkp_h_1:msas_u_1)

# Now let's create a more amenable dataset 

d <- d %>% 
  mutate(id = 1:n()) %>% 
  select(id, everything()) 

# Rename all variables 

d <- d %>%  
  rename_all(
    funs(str_remove(., "_1"))
    ) %>% 
  mutate_if(is.character, as.numeric)

```

At this point, it is also useful to put the data into a long format for multilevel analysis and in order to create certain visualizations. 

```{r}

# Reshape harmful perceptions from wide to long 

d_harm_long <- d %>% 
  select(id, contains("_h")) %>% 
  gather(key = "scenario", value = "harm", 2:26) %>% 
  mutate(scenario = str_remove(scenario, "_h$"))

# Reshape immoral perceptions from wide to long 

d_imm_long <- d %>% 
  select(id, contains("_i")) %>% 
  gather(key = "scenario", value = "immoral", 2:26) %>% 
  mutate(scenario = str_remove(scenario, "_i$"))

# Reshape unexpected perceptions from wide to long 

d_un_long <- d %>% 
select(id, contains("_u")) %>% 
  gather(key = "scenario", value = "unexpected", 2:26) %>% 
  mutate(scenario = str_remove(scenario, "_u$"))

# Now join them 

d_long <- d_harm_long %>% 
  mutate(immoral = d_imm_long$immoral, 
         unexpected = d_un_long$unexpected)
```

At this point it would be useful to create a column for each moral foundation and to add the deflection to each event. 

```{r}

# Load dataset with the deflections 

selected_events <- read_csv("selected_events.csv")

# Create column for abreviations 

events <- c("pch", 
            "pkp", 
            "tmdp", 
            "adr", 
            "elb", 
            "acr", 
            "jbd", 
            "scc", 
            "ddf", 
            "git", 
            "idb", 
            "ayc", 
            "sip", 
            "ecc", 
            "ccr", 
            "mbr", 
            "msn", 
            "mbw", 
            "php", 
            "mmc", 
            "mmls", 
            "msas", 
            "mpsa", 
            "whh", 
            "thls")

# Create new column 

events_def <- selected_events %>% 
  mutate(scenario = events) %>% 
  select(scenario, def, type)

# Add to long dataset 

d_long <- d_long %>% 
  left_join(events_def, by = "scenario")
```

Now I am going to create a summary table for each question 

```{r}

# create a summary table for the values of each question 

gd <- d %>% 
  select(-id) %>% 
  gather(key = "scenario", 
         value = "score") %>% 
  group_by(scenario) %>% 
  summarise_all(funs(avg = mean(.), maximum = max(.), minimum = min(.), st_dv = sd(.))) %>% 
  mutate(ci_lower = avg - 1.96*st_dv, 
         ci_upper = avg + 1.96*st_dv) 

gd <- gd %>% 
  mutate(scenario = str_replace(scenario, "acr", "athlete_cheats_rival"), 
         scenario = str_replace(scenario, "adr", "athlete_deceives_referee"), 
         scenario = str_replace(scenario, "ayc", "athlete_yells_at_coach"), 
         scenario = str_replace(scenario, "ccr", "coach_cheers_rival"), 
         scenario = str_replace(scenario, "ddf", "daughter_disobeys_father"),
         scenario = str_replace(scenario, "ecc", "employee_conspires_with_comp."),
         scenario = str_replace(scenario, "elb", "employee_lies_to_boss"), 
         scenario = str_replace(scenario, "git", "girl_interrupts_teacher"),
         scenario = str_replace(scenario, "idb", "intern_disobeys_boss"), 
         scenario = str_replace(scenario, "jbd", "judge_befriends_defendant"), 
         scenario = str_replace(scenario, "mbr", "man_betrays_relative"), 
         scenario = str_replace(scenario, "mbw", "man_betrays_wife"), 
         scenario = str_replace(scenario, "mmc", "man_marries_cousin"),
         scenario = str_replace(scenario, "mmls", "man_makes_love_sister"), 
         scenario = str_replace(scenario, "mpsa", "married_has_sex_adulterer"), 
         scenario = str_replace(scenario, "msas", "mother_sexually_arouses_son"), 
         scenario = str_replace(scenario, "msn", "mayor_slanders_neighbor"), 
         scenario = str_replace(scenario, "pch", "person_hurts_child"), 
         scenario = str_replace(scenario, "php", "person_hires_prostitute"), 
         scenario = str_replace(scenario, "pkp", "person_kills_person"),
         scenario = str_replace(scenario, "scc", "student_cheats_classmate"), 
         scenario = str_replace(scenario, "sip", "student_insults_professor"), 
         scenario = str_replace(scenario, "thls", "teacher_hits_lazy_student"), 
         scenario = str_replace(scenario, "whh", "wife_hits_husband"), 
         scenario = str_replace(scenario, "tmdp", "teenager_mocks_disabled"))
  
```


Now, we are in a position to start making some summary plots 

```{r}

# Keep only harm questions 

gdh <- gd[str_which(gd$scenario, "_h$"),]

# Plot harm questions 

gdh %>% 
  mutate(scenario = str_remove(scenario, "_h$")) %>% 
  ggplot(aes(x = reorder(scenario, avg), y = avg)) + 
  geom_point(col = "darkred", fill = "darkred") + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  coord_flip() + 
  labs(x = "", 
       y = "Ratings", 
       title = "Perceptions of Harmfulness")

# Now lets look at immorality 

gdi <- gd[str_which(gd$scenario, "_i$"),]

# Plot immorality

gdi %>% 
  mutate(scenario = str_remove(scenario, "_i$")) %>% 
  ggplot(aes(x = reorder(scenario, avg), y = avg)) + 
  geom_point(col = "darkred", fill = "darkred") + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  coord_flip() + 
  labs(x = "", 
       y = "Ratings", 
       title = "Perceptions of Immorality")

# Lastly let's look at unexpectedness 

gdu <- gd[str_which(gd$scenario, "_u$"),]

# Plot Unexpectedness

gdu %>% 
  mutate(scenario = str_remove(scenario, "_u$")) %>% 
  ggplot(aes(x = reorder(scenario, avg), y = avg)) + 
  geom_point(col = "darkred", fill = "darkred") + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  coord_flip() + 
  labs(x = "", 
       y = "Ratings", 
       title = "Perceptions of Unexpectedness")

```


