---
title: "Moral Scenarios and Affect Control Theory"
author: "Nicolas Restrepo"
date: "September 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```


```{r}

library(tidyverse)
library(readxl)
library(stringr)
theme_set(theme_bw())
```

## Preliminary Analysis 

Load in the data 

```{r}

# Load the dataset 

moral_events <- read_csv("moral_events.csv")

```

Are there types of moral scenarios that are more deflecting than others? 

```{r}

mean_def <- moral_events %>% 
  group_by(type) %>% 
  summarize( mean_deflection = mean(def), 
             sd_deflection = sd(def), 
             upper = mean_deflection + 1.96*sd_deflection, 
             lower = mean_deflection - 1.96*sd_deflection)

mean_def %>% 
  ggplot(aes(x = reorder(type, mean_deflection), y = mean_deflection, ymin = lower, ymax = upper)) + 
  geom_point() + 
  geom_pointrange() + 
  coord_flip() + 
  labs( title = "Average deflection in each category", 
        x = "", 
        y = "Average Deflection")
```

There is not much that can be inferred from this. It is just immediately interesting how deflection is a very good predictor of the "wrongness" of harm but a terrible predictor of the "wrongness" of purity. 

What about the situations themselves? 

```{r}

moral_events <- moral_events %>% 
  mutate(event = paste(actor, behavior, object, sep = " ")) %>% 
  filter(type != "liberty")

moral_events %>% 
  ggplot(aes(x = reorder(event, def), y = def)) + 
  geom_point() + 
  coord_flip() + 
  facet_grid(type ~ . , scales = "free_y", space = "free") +
  labs(title = "Deflection of the Moral Scenarios", 
       y = " ", 
       x = "") + 
  theme(legend.position = "none") +
  theme_minimal(base_size = 10) 
```

This is quite incredible. Deflection is a good measure of harm and authority, because those features are built into the equations. However, it cannot account for the wrongness of situations such as incest or pedophilia. 

