---
title: "Reaction Time - Pretest"
author: "Nicolas Restrepo"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analyzing Data from Pre-Test

```{r, message=FALSE}
library(tidyverse)
library(kableExtra)
library(naniar)
theme_set(theme_light())
```

I'm going to begin by loading in the data. 

```{r, message = FALSE}
# Load in the data 
rt_pretest <- read_csv("data/rt_pretest.csv")
```

As a first step I am going to check how people did in the attention checks. 

```{r}
# Delete unnecessary rows
rt_pretest <- rt_pretest[-(1:2),]

# How many people passed the attention checks?
rt_pretest %>% 
  select(starts_with("att")) %>% 
  mutate(passed = if_else(att_1 == "Harmless (K)" & att_2 == "Immoral (J)" & att_3 == "Harmful (J)", 1, 0)) %>% 
  group_by(passed) %>% 
  summarise(n())
```

Everyone passed! 

Now let's see whether we can get reaction time. 

```{r}
# Put data in long format
rt_pretest_times <- rt_pretest %>% 
  select(ends_with("Page Submit")) %>% 
  mutate(ids = 1:20) %>% 
  pivot_longer(1:50, names_to = "scenario", values_to = "reaction_time") %>% 
  mutate(scenario = str_remove(scenario, "_t_Page Submit$"))

# Get summary statistics 
gd <- rt_pretest_times %>% 
  mutate(reaction_time = as.integer(reaction_time)) %>% 
  group_by(scenario) %>% 
  summarise(avg = mean(reaction_time), 
            st_dv = sd(reaction_time)) %>% 
  mutate(ci_lower = avg - 1.96*st_dv, 
         ci_upper = avg + 1.96*st_dv)

# Now let's plot harm 

# Keep only harm questions 

gdh <- gd[str_which(gd$scenario, "h_"),]

# Plot harm questions 

gdh %>% 
  mutate(scenario = str_remove(scenario, "h_")) %>% 
  ggplot(aes(x = reorder(scenario, avg), y = avg)) + 
  geom_point(col = "darkred", fill = "darkred") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  coord_flip() + 
  labs(x = "", 
       y = "RT", 
       title = "Reaction Time - Harmfulness", 
       caption = "Circles = Mean")

# Get only Immorality questions 
gdh <- gd[str_which(gd$scenario, "i_"),]

# Plot harm questions 

gdh %>% 
  mutate(scenario = str_remove(scenario, "i_")) %>% 
  ggplot(aes(x = reorder(scenario, avg), y = avg)) + 
  geom_point(col = "darkred", fill = "darkred") + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  coord_flip() + 
  labs(x = "", 
       y = "RT", 
       title = "Reaction Time - Immorality",
       caption = "Circles = Mean")
```

We only have 20 cases so the confidence intervals will be large. The highest mean is "married person has sex with an adulterer", which is probably a result of the length of the sentence. This is something we will need to control for in the final analysis. The fact that the "prototype" has the lowest response time in both immorality and harmfulness is really encouraging. 

```{r}

rt_pretest_times %>% 
  mutate(reaction_time = as.integer(reaction_time)) %>% 
  summarise(avg = mean(reaction_time), 
            stv = sd(reaction_time)) %>% 
  mutate(ci_lower = avg - 1.96*stv, 
         ci_upper = avg + 1.96*stv) %>% 
  kable(.) %>% 
  kable_styling()

sum(rt_pretest_times$reaction_time < 1)
sum(rt_pretest_times$reaction_time > 5.46)
sum(rt_pretest_times$reaction_time > 9)

  
```

It is also encouraging that only one of the responses (total = 1000) was below 1 second. This means that no one just clicked their way through the survey. Also, the upper bound of the 95% confidence interval of reaction times is 5.46. 6.9% of the responses were longer than this threshold. This indicates that most people took a "reasonable" time to answer questions. Now, we will have to think more about outliers later but it is encouraging that - in general - respondents were neither just clicking through the questions nor taking too long. 


